maASlin2

### 2019-12-17 scripts/27_maASlin2.Rmd
update

```{r}
# get the MRN of my target samples 
link <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 11252019 wo Censor .xlsx',
                    sheet = 4) %>% 
  dplyr::select(pt_id, MRN)

@@

# sub selecting the vars I'm interested in 
pheno <- phe_M %>% 
  dplyr::select(MRN,
                sampleid = Baseline_Sample_ID,
                CR, 
                ORR, 
                Toxicity, 
                Cytokine_Release_Syndrome,
                ICANS,
                Antibiotics_4_weeks_prior_to_CAR_T_cells,
                #Antibiotics_4_weeks_post_CAR_T_cells,
                Disease_Class,
                Key__Patient_Disease_Status_at_CAR
                )
```

@@

```{r}
# get the clinical data info  from ying pulled data
load('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/YingRCode/ctc_melody.RData')
```

@@

# the three ILs, select the highest value if there are multiple 
pheno <- pheno %>% 
  # get patient info such as age and sex (using MRN)
  left_join(ctc.patients %>% 
              dplyr::select(MRN, sex, dob), by = 'MRN') %>% 
  # get the hightest ferritin value 
  left_join(ctc.ferritin %>% 
              dplyr::select(MRN, result_text) %>% 
              group_by(MRN) %>% 
              summarise(ferritin = max(result_text)), by = 'MRN') %>% 
  left_join(ctc.crp %>% 
              dplyr::select(MRN, result_text) %>% 
              mutate(result_text = if_else(str_detect(result_text, '^<'), '0', result_text),
                     result_text = as.numeric(result_text)) %>% 
              group_by(MRN) %>% 
              summarise(crp  = max(result_text, na.rm = T)), by = 'MRN') %>% 
  left_join(IL_table, by  ='MRN') %>% 
  left_join(lines, by = 'MRN')

@@

```{r pheno_final}
# now change the columns to different type: dbl or factor
pheno_final  <- pheno %>%  
  # change the ferritin to dbl
  mutate(ferritin = as.numeric(ferritin),
         Key__Patient_Disease_Status_at_CAR = as.character(Key__Patient_Disease_Status_at_CAR)) %>% 
  # change all the pheno vars to factors
  mutate_if(is.character, as.factor) %>% 
  mutate(MRN = as.character(MRN),
         sampleid = as.character(sampleid)) %>% 
  #mutate(IL_6 = log(IL_6),
  #       IL_10 = log(IL_10)) %>% 
  column_to_rownames('sampleid')

@@

  round(cor.test(pheno_final[,Row[['V1']]],  pheno_final[,Row[['V2']]],method = 'spearman', 
                           exact = F)$estimate, 2)
  #chisq.test(phe_fil)$p.value
}) %>% 
  as.data.frame() %>% 
  bind_cols(combi) %>% 
  mutate(phe_var = str_glue('{V1}__{V2}')) %>% 
  dplyr::select(-V1, -V2)

@@

```{r}

fit_data <- Maaslin2(
    pw, pheno_final, '../data/maASlin2/pw', 
    normalization  = 'NONE', 
    transform = "LOG",  
    cores = 4, 
    plot_heatmap = T,
    min_prevalence = 0.5,
    fixed_effects = c('Toxicity', 
                      'CR',
                      'Disease_Class',
                      'Key__Patient_Disease_Status_at_CAR',
                      'Antibiotics_4_weeks_prior_to_CAR_T_cells', 
                      'ferritin',
                      'crp',
                      'IL_1_beta',
                      'IL_10',
                      'IL_6',
                      'Prior_lines_of_therapy'),
    random_effects = c('sex',
                       'age'),
    standardize = FALSE,
    max_significance = 0.1)

@@

>>>>

```{r}
# get the MRN of my target samples 
link <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 12122019.xlsx',
                    sheet = 4) %>% 
  dplyr::select(pt_id, 
                MRN,
                disease_sta = `Key: Patient Disease Status at CAR`)

@@

# sub selecting the vars I'm interested in 
pheno <- phe_M %>% 
  dplyr::select(MRN,
                sampleid = Baseline_Sample_ID,
                CR, 
                ORR, 
                Toxicity, 
                Cytokine_Release_Syndrome,
                ICANS,
                Antibiotics_4_weeks_post_CAR_T_cells,
                Disease_Class,
                disease_sta
                )
```

@@

```{r}
# get the clinical data info  from ying pulled data
load('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/YingRCode/ctc_melody.RData')

# load the table I made for the abx exposure
abx <- read_csv('../data/cleaned_pheno/shotgun_cohort_abx_4w_baseline_expo_table.csv')
```

@@

# the three ILs, select the highest value if there are multiple 
pheno <- pheno %>% 
  # get patient info such as age and sex (using MRN)
  left_join(ctc.patients %>% 
              dplyr::select(MRN, sex, dob), by = 'MRN') %>% 
  # get the hightest ferritin value 
  left_join(ctc.ferritin %>% 
              dplyr::select(MRN, result_text) %>% 
              group_by(MRN) %>% 
              summarise(ferritin = max(result_text)), by = 'MRN') %>% 
  left_join(ctc.crp %>% 
              dplyr::select(MRN, result_text) %>% 
              mutate(result_text = if_else(str_detect(result_text, '^<'), '0', result_text),
                     result_text = as.numeric(result_text)) %>% 
              group_by(MRN) %>% 
              summarise(crp  = max(result_text, na.rm = T)), by = 'MRN') %>% 
  left_join(IL_table, by  ='MRN') %>% 
  left_join(lines, by = 'MRN') %>% 
  left_join(abx, by = 'MRN')

@@

pheno %>% 
  write_csv('../data/cleaned_pheno/final_pheno.csv')
```

```{r pheno_final}
# now change the columns to different type: dbl or factor
pheno_final  <- pheno %>%  
  # change the ferritin to dbl
  mutate(ferritin = as.numeric(ferritin),
         disease_sta = as.character(disease_sta)) %>% 
  # change all the pheno vars to factors
  mutate_if(is.character, as.factor) %>% 
  mutate(MRN = as.character(MRN),
         sampleid = as.character(sampleid)) %>% 
  #mutate(IL_6 = log(IL_6),
  #       IL_10 = log(IL_10)) %>% 
  column_to_rownames('sampleid')

@@

  round(cor.test(pheno_final[,Row[['V1']]],  pheno_final[,Row[['V2']]],method = 'spearman', 
                           exact = F)$estimate, 2)
  #chisq.test(phe_fil)$p.value
}) %>% 
  as.data.frame() %>% 
  bind_cols(combi) %>% 
  mutate(phe_var = str_glue('{V1}__{V2}')) %>% 
  dplyr::select(-V1, -V2) 
```

@@

```{r}
# a look at the final pheno data
colnames(pheno_final)

cor.test(pheno_final$crp, pheno_final$IL_6, method = 'pearson')
```

```{r}
# remove everything already in this folder
#system('rm ../data/maASlin2/pw/*')

fit_data <- Maaslin2(
    pw, pheno_final, '../data/maASlin2/pw/onlyFixed', 
    normalization  = 'NONE', 
    transform = "LOG",  
    cores = 4, 
    plot_heatmap = F,
    min_abundance = 0.05,
    min_prevalence = 0.1,
    fixed_effects = c('Toxicity', 
                      'CR',
                      'Disease_Class',
                      'disease_sta',
                      'abx_expo', 
                      'ferritin',
                      'crp',
                      'IL_1_beta',
                      #'IL_10',
                      'IL_6',
                      'Prior_lines_of_therapy',
                      'Antibiotics_4_weeks_post_CAR_T_cells'),
    # random_effects = c('sex',
    #                   'age'),
    standardize = FALSE,
    max_significance = 0.25)

@@

```{r}
system('rm ../data/maASlin2/pw/both/*')

fit_data <- Maaslin2(
    pw, pheno_final, '../data/maASlin2/pw/both', 
    normalization  = 'NONE', 
    transform = "LOG",  
    cores = 4, 
    plot_heatmap = F,
    min_abundance = 0.05,
    min_prevalence = 0.1,
    fixed_effects = c('Toxicity', 
                      'CR',
                      'Disease_Class',
                      'disease_sta',
                      'abx_expo', 
                      'ferritin',
                      'crp',
                      #'IL_1_beta',
                      #'IL_10',
                      #'IL_6',
                      'Prior_lines_of_therapy',
                      'Antibiotics_4_weeks_post_CAR_T_cells'),
    random_effects = c('sex',
                      'age'),
    standardize = FALSE,
    max_significance = 0.25)
```
